---
title: "Writing Functions"
output: github_document
date: "2023-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
```


Set the seed for reproducibility.

```{r}
set.seed(12345)
```

### Z score function

Z scores subtract the mean and divid by the sd.

```{r}
x_vec = rnorm(20, mean = 5, sd = 0.3)
```

Compute Z scores for `x_vec`.

```{r}
(x_vec - mean(x_vec)) / sd(x_vec)
```

Write a function to do this!

```{r}
z_score = function(x) {
  
  if (!is.numeric(x)) {
    
    stop("Argument should be numbers")
    
  }  else if (length(x) < 2) {
    
    stop("You need at least 2 numbers to get z scores")
    
  }
  
  z = (x-mean(x)) / sd(x) 
 
  z
 
}
```

Check that this works.

```{r}
z_score(x = x_vec)

z_score(x = rnorm(10, mean = 5))
```

Keep checking. 

```{r}
#z_score(x = 3)

#z_score(c("my", "name", "is"))

#z_score(c(TRUE, TRUE, FALSE, TRUE))

#z_score(iris)
```

### Multiple outputs

Write a function that returns the mean and sd from a sample of numbers.

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    
    stop("Argument should be numbers")
    
  }  else if (length(x) < 2) {
    
    stop("You need at least 2 numbers to get z scores")
    
  }
  
  mean_x = mean(x)
  
  sd_x = sd(x)
  
  tibble(
    mean = mean_x,
    sd = sd_x
  )
}
```

Double check 

```{r}
mean_and_sd(x_vec)
```

### Multiple inputs

Start getting means and sds
```{r}
x_vec = rnorm(n = 30, mean = 5, sd = 0.5)

tibble(
  mean = mean(x_vec),
  sd = sd(x_vec)
)
```

Let's write a function that uses `n`, a true mean, and true SD as inputs.

```{r}
sim_mean_sd = function(n_obs, mu, sigma) {
  
  x_vec = rnorm(n = n_obs, mean = mu, sd = sigma)

  tibble(
    mean = mean(x_vec),
    sd = sd(x_vec)
)
}

sim_mean_sd(n_obs = 30, mu = 50, sigma = 12.3)
```

Think about default values and positional matching. 


